### 2016-03-12 paper.tex
init

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\begin{document}

\title{Hypothetical Problems concerning the Theory of Relativity on Cryptographic Currency Implementations}
\author
{\IEEEauthorblockN{Abrahim Ladha}
\IEEEauthorblockA{Computer Security\\
Armstrong State University\\
abrahimladha@protonmail.ch}
}
\maketitle

\begin{abstract}
this is my abstract section
\end{abstract}

\section{What is bitcoin?}
\textbf{History} Bitcoin was invented by someone using the psuedonym Satoshi Nakamoto. In his original whitepaper titled "Bitcoin: A Peer-to-Peer Electronic Cash System", he introduces the concept of a decentralized and crytpographically secured monetary system. He also covers topics on proof-of-work, transactions, mining, privacy, and attacks on the network.\cite{nakamoto} It was implemented as free software and released in January 2009. Unlike gold, bitcoin has no attachment to any sort of industry, so the price day to day price fluctuation is really based on nothing but speculation. It has been as high as 1000 USD per bitcoin.\cite{1000usd}

\textbf{Units} There can only ever exist a maximum of 21 million bitcoins. Each bitcoin can be broken up into a hundred million pieces, much like a single US Dollar can be broken up into one hundred pennies. The smallest unit of bitcoin is called a satoshi.

\textbf{Ownership} Since bitcoin has no central authority, transactions are made user to user. A user can only send bitcoins to another user if they can digitally sign the transaction with their private key. Without the private key, a malicious user cannot sign the transaction and the coins cannot be spent.

\textbf{transactions} Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\textbf{mining} Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\section{The Theory of Relativity}
\textbf{history} Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\textbf{lorentz equations} Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\textbf{time travel} Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\textbf{lizard problem} Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\bibliographystyle{IEEEtran}
\begin{thebibliography}{9}

\bibitem{nakamoto}
	N. Satoshi. 
	(2008)
	\textit{Bitcoin: A Peer-to-Peer Electronic Cash System}
	[Whitepaper]
	Available: https://bitcoin.org/bitcoin.pdf

\bibitem{1000usd}
	M. Farrell. 
	(2013).	
	\textit{What Bubble? Bitcoin Tops \$1000}
	[Online].
	Available: https://blogs.wsj.com/moneybeat/2013/11/27/what-bubble-bitcoin-tops-1000/
\bibitem{template}
	J. K. Author. 
	(year).	
	\textit{Title}
	(edition)
	[Type of medium].
	Available: https://www.example.com
\end{thebibliography}
\end{document}